Generating gaits for physical quadruped robots: evolved neural networks vs. local parameterized search

### 2011-04-12 paper_ecal/hyperNeatMethod.tex
changes to hyperneat section

The ANN configuration follows previous studies that evolved quadruped
gaits with HyperNEAT in simulation~\cite{clune2011performance,
  clune2009evolving}, but was adapted to accommodate the physical robot used in this paper. Specifically, the ANNs consists of three $3 \times 4$
Cartesian grids of nodes forming input, hidden, and output
layers (\figref{SpiderANN.jpg}). Adjacent layers were allowed to be completely connected, meaning that there
could be $(3 \times 4)^2= 288$ links in each ANN (although evolution can set weights to 0, functionally eliminating the connection). The inputs to the
substrate were the current \emph{commanded} angles of each of the 9 joints of the robot and a sine and cosine wave (to facilitate the
production of periodic behaviors).  The sine and cosine waves had a period of about half a second.

@@

The outputs of the substrate at each time step were nine numbers in
the range $[-1,1]$, which were scaled according to the allowable
ranges for each of the nine motors and then commanded
the positions for each motor.  Occasionally HyperNEAT would produce networks that
exhibited rapid oscillatory behaviors, switching from extreme negative to extreme positive numbers each time step.  This resulted in motor
commands to alternate extremes every 25ms (given the command rate of
40Hz), which tended to damage and overheat the motors.  To ameliorate
this problem, we requested four times as many commanded
positions from HyperNEAT ANN's and averaged over four commands at a time to
obtain the actual gait $g(t)$.  This solution worked well and did not restrict the expressiveness of HyperNEAT.

%\editbox{are we mentioning that we gave HyperNEAT the center joint
%  here or earlier?}

@@

As with the parameterized methods, three runs of HyperNEAT
were performed. The population size for HyperNEAT was 9 and runs
lasted 20 generations. These numbers are small, but were necessarily
constrained given how much time it took to conduct evolution directly
on a real robot. The remaining parameters were identical to Clune et
al.~\cite{clune2011performance}.

>>>>

The ANN configuration follows previous studies that evolved quadruped
gaits with HyperNEAT in simulation~\cite{clune2011performance,
  clune2009evolving}, but was adapted to accommodate the physical robot used in this paper. Specifically, the ANN has a fixed topology (i.e. the number of nodes does not evolve) that consists of three $3 \times 4$
Cartesian grids of nodes forming input, hidden, and output
layers (\figref{SpiderANN.jpg}). Adjacent layers were allowed to be completely connected, meaning that there
could be $(3 \times 4)^2= 288$ links in each ANN (although evolution can set weights to 0, functionally eliminating the connection). The inputs to the
substrate were the current \emph{commanded} angles of each of the 9 joints of the robot and a sine and cosine wave (to facilitate the
production of periodic behaviors).  The sine and cosine waves had a period of about half a second.

@@

The outputs of the substrate at each time step were nine numbers in
the range $[-1,1]$, which were scaled according to the allowable
ranges for each of the nine motors and then commanded
the positions for each motor.  Occasionally HyperNEAT would produce networks that
exhibited rapid oscillatory behaviors, switching from extreme negative to extreme positive numbers each time step.  This resulted in motor
commands to alternate extremes every 25ms (given the command rate of
40Hz), which tended to damage and overheat the motors.  To ameliorate
this problem, we requested four times as many commanded
positions from HyperNEAT ANN's and averaged over four commands at a time to
obtain the actual gait $g(t)$.  This solution worked well and did not restrict the expressiveness of HyperNEAT.
%\editbox{are we mentioning that we gave HyperNEAT the center joint
%  here or earlier?}

@@

As with the parameterized methods, three runs of HyperNEAT
were performed. Runs lasted 20 generations and the population size was 9 (which allowed a bare minimum of diversity within and between NEAT species).  These numbers are small, but were necessarily
constrained given how much time it took to conduct evolution directly
on a real robot. The remaining parameters were identical to Clune et
al.~(2011).
